---
title: "Agricultural Analysis"
format: 
  html:
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(plotly)
library(reactable)
library(DT)
library(htmlwidgets)
library(crosstalk)
library(htmltools)

# Custom color palette
custom_colors <- list(
  primary = "#2E8B57",
  secondary = "#20B2AA", 
  success = "#28a745",
  warning = "#ffc107",
  danger = "#dc3545",
  info = "#17a2b8",
  light = "#f8f9fa",
  dark = "#343a40",
  gradient = c("#2E8B57", "#20B2AA", "#66CDAA", "#98FB98", "#90EE90")
)

# Helper functions
format_percent <- function(x, digits = 1) {
  paste0(round(x * 100, digits), "%")
}

format_number <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", scientific = FALSE)
}

# Load and process data
df_raw <- readRDS("myapp/df.rds")

df_processed <- df_raw %>%
  mutate(across(everything(), ~ na_if(., "---"))) %>%
  mutate(across(where(is.list), ~ sapply(., function(x) if(length(x) > 0) x[1] else NA))) %>%
  mutate(
    completed_time = ymd_hms(completed_time),
    started_time = ymd_hms(started_time)
  ) %>%
  mutate(
    date = as.Date(completed_time),
    week = floor_date(date, "week"),
    month = floor_date(date, "month"),
    hour_started = hour(started_time),
    day_of_week = wday(date, label = TRUE, week_start = 1),
    is_weekend = day_of_week %in% c("Sat", "Sun"),
    duration_minutes = as.numeric(difftime(completed_time, started_time, units = "mins")),
    is_night_survey = hour_started >= 19 | hour_started < 6,
    is_short_survey = duration_minutes <= 5,
    is_long_survey = duration_minutes >= 60
  ) %>%
  mutate(across(c(starts_with("ps_num_planted_"), starts_with("num_surv_")), as.numeric))

df_completed <- df_processed %>% filter(consent == 1)

# Species analysis
species_data <- df_completed %>%
  select(starts_with("ps_num_planted_"), starts_with("num_surv_")) %>%
  pivot_longer(everything(), names_to = "key", values_to = "value") %>%
  mutate(
    type = ifelse(str_detect(key, "ps_num_planted"), "planted", "survived"),
    species = str_remove_all(key, "ps_num_planted_|num_surv_")
  ) %>%
  group_by(species, type) %>%
  summarise(total = sum(value, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = type, values_from = total) %>%
  mutate(
    planted = ifelse(is.na(planted), 0, planted),
    survived = ifelse(is.na(survived), 0, survived),
    survival_rate = ifelse(planted > 0, (survived / planted) * 100, 0),
    loss_count = planted - survived,
    loss_rate = ifelse(planted > 0, (loss_count / planted) * 100, 0)
  ) %>%
  filter(planted > 0) %>%
  arrange(desc(survival_rate))

# Create shared data for crosstalk
shared_species_data <- SharedData$new(species_data)
```

::: {.filter-section}
## 🔍 Species Analysis Filters

::: {.interactive-controls}
```{r species_filters}
#| echo: false

# Species filter
filter_select("species", "Species", shared_species_data, ~species, multiple = TRUE)

# Survival rate filter
filter_slider("survival_rate", "Survival Rate (%)", shared_species_data, ~survival_rate)

# Planted count filter
filter_slider("planted", "Trees Planted", shared_species_data, ~planted)
```
:::
:::

## 🌱 Species Performance Overview

::: {.dashboard-grid}
::: {.metric-card}
::: {.metric-label}
Total Species Tracked
:::
::: {.metric-value}
`r nrow(species_data)`
:::
:::

::: {.metric-card}
::: {.metric-label}
Total Trees Planted
:::
::: {.metric-value}
`r format_number(sum(species_data$planted, na.rm = TRUE))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Total Trees Survived
:::
::: {.metric-value}
`r format_number(sum(species_data$survived, na.rm = TRUE))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Overall Survival Rate
:::
::: {.metric-value}
`r format_percent(sum(species_data$survived, na.rm = TRUE) / sum(species_data$planted, na.rm = TRUE))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Best Performing Species
:::
::: {.metric-value}
`r species_data$species[1]`
:::
:::
:::

## 🌱 Tree Survival Rates by Species

::: {.plot-container}
```{r survival_rates_plot}
#| echo: false

p1 <- plot_ly(species_data, 
        x = ~reorder(species, survival_rate), 
        y = ~survival_rate, 
        type = 'bar', 
        marker = list(
          color = ~survival_rate, 
          colorscale = list(c(0, "#ffcdd2"), c(1, custom_colors$success)),
          showscale = TRUE,
          colorbar = list(title = "Survival Rate (%)")
        ),
        hovertemplate = "Species: %{x}<br>Survival Rate: %{y:.1f}%<br>Planted: %{customdata[0]}<br>Survived: %{customdata[1]}<extra></extra>",
        customdata = ~cbind(planted, survived)) %>%
  layout(
    title = "Tree Survival Rates by Species",
    xaxis = list(title = "Species", tickangle = -45),
    yaxis = list(title = "Survival Rate (%)", range = c(0, 100)),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p1
```
:::

## 📊 Detailed Species Performance Table

::: {.table-container}
```{r species_table}
#| echo: false

reactable(
  shared_species_data,
  columns = list(
    species = colDef(
      name = "Species", 
      minWidth = 150,
      cell = function(value) {
        div(style = "font-weight: bold; color: #2E8B57;", value)
      }
    ),
    planted = colDef(
      name = "Trees Planted", 
      align = "center",
      format = colFormat(separators = TRUE)
    ),
    survived = colDef(
      name = "Trees Survived", 
      align = "center",
      format = colFormat(separators = TRUE)
    ),
    loss_count = colDef(
      name = "Trees Lost", 
      align = "center",
      format = colFormat(separators = TRUE),
      cell = function(value) {
        if (value > 0) {
          div(style = "color: #dc3545; font-weight: bold;", format(value, big.mark = ","))
        } else {
          div(style = "color: #28a745;", "0")
        }
      }
    ),
    survival_rate = colDef(
      name = "Survival Rate (%)",
      align = "center",
      format = colFormat(digits = 1),
      cell = function(value) {
        color <- if (value >= 80) {
          "#A5D6A7"
        } else if (value >= 60) {
          "#FFF9C4"
        } else {
          "#FFCDD2"
        }
        div(
          style = paste("background:", color, "; padding: 4px; border-radius: 4px; font-weight: bold;"),
          paste0(round(value, 1), "%")
        )
      }
    ),
    loss_rate = colDef(
      name = "Loss Rate (%)",
      align = "center",
      format = colFormat(digits = 1),
      cell = function(value) {
        color <- if (value <= 20) {
          "#A5D6A7"
        } else if (value <= 40) {
          "#FFF9C4"
        } else {
          "#FFCDD2"
        }
        div(
          style = paste("background:", color, "; padding: 4px; border-radius: 4px; font-weight: bold;"),
          paste0(round(value, 1), "%")
        )
      }
    )
  ),
  defaultPageSize = 10, 
  searchable = TRUE, 
  highlight = TRUE, 
  bordered = TRUE, 
  striped = TRUE,
  theme = reactableTheme(
    headerStyle = list(backgroundColor = "#f8f9fa", fontWeight = "bold"),
    cellStyle = list(fontSize = "0.9em")
  )
)
```
:::

## 📈 Planted vs Survived Comparison

::: {.plot-container}
```{r planted_vs_survived}
#| echo: false

p2 <- plot_ly(species_data, 
        x = ~planted, 
        y = ~survived, 
        type = 'scatter',
        mode = 'markers',
        marker = list(
          size = 12,
          color = ~survival_rate,
          colorscale = list(c(0, "#ffcdd2"), c(1, custom_colors$success)),
          showscale = TRUE,
          colorbar = list(title = "Survival Rate (%)")
        ),
        text = ~species,
        hovertemplate = "Species: %{text}<br>Planted: %{x}<br>Survived: %{y}<br>Survival Rate: %{marker.color:.1f}%<extra></extra>") %>%
  layout(
    title = "Trees Planted vs Trees Survived",
    xaxis = list(title = "Trees Planted"),
    yaxis = list(title = "Trees Survived"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    shapes = list(
      list(
        type = "line",
        x0 = 0, x1 = max(species_data$planted),
        y0 = 0, y1 = max(species_data$planted),
        line = list(color = "gray", dash = "dash", width = 2)
      )
    ),
    annotations = list(
      list(
        x = max(species_data$planted) * 0.8,
        y = max(species_data$planted) * 0.9,
        text = "Perfect Survival (100%)",
        showarrow = FALSE,
        font = list(color = "gray")
      )
    )
  )

p2
```
:::

## 🏆 Top and Bottom Performers

::: {.plot-container}
```{r top_bottom_performers}
#| echo: false

# Get top 5 and bottom 5 performers
top_performers <- species_data %>%
  top_n(5, survival_rate) %>%
  mutate(category = "Top 5")

bottom_performers <- species_data %>%
  top_n(-5, survival_rate) %>%
  mutate(category = "Bottom 5")

performers_data <- bind_rows(top_performers, bottom_performers)

p3 <- plot_ly(performers_data, 
        x = ~reorder(species, survival_rate), 
        y = ~survival_rate,
        color = ~category,
        colors = c("Top 5" = custom_colors$success, "Bottom 5" = custom_colors$danger),
        type = 'bar',
        hovertemplate = "Species: %{x}<br>Survival Rate: %{y:.1f}%<br>Category: %{fullData.name}<extra></extra>") %>%
  layout(
    title = "Top 5 vs Bottom 5 Performing Species",
    xaxis = list(title = "Species", tickangle = -45),
    yaxis = list(title = "Survival Rate (%)", range = c(0, 100)),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    legend = list(orientation = 'h', y = 1.1, x = 0.5, xanchor = 'center'),
    margin = list(b = 120)
  )

p3
```
:::

## 🌍 Species Distribution Analysis

::: {.plot-container}
```{r species_distribution}
#| echo: false

# Pie chart of planted trees by species
p4 <- plot_ly(species_data, 
        labels = ~species, 
        values = ~planted, 
        type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hovertemplate = "Species: %{label}<br>Trees Planted: %{value}<br>Percentage: %{percent}<extra></extra>",
        marker = list(
          colors = custom_colors$gradient,
          line = list(color = '#FFFFFF', width = 2)
        )) %>%
  layout(
    title = "Distribution of Planted Trees by Species",
    paper_bgcolor = "transparent",
    showlegend = FALSE
  )

p4
```
:::

## 📊 Survival Rate Categories

::: {.plot-container}
```{r survival_categories}
#| echo: false

# Categorize species by survival rate
species_categories <- species_data %>%
  mutate(
    category = case_when(
      survival_rate >= 80 ~ "Excellent (≥80%)",
      survival_rate >= 60 ~ "Good (60-79%)",
      survival_rate >= 40 ~ "Fair (40-59%)",
      TRUE ~ "Poor (<40%)"
    )
  ) %>%
  count(category) %>%
  mutate(
    category = factor(category, levels = c("Excellent (≥80%)", "Good (60-79%)", "Fair (40-59%)", "Poor (<40%)"))
  )

p5 <- plot_ly(species_categories, 
        x = ~category, 
        y = ~n, 
        type = 'bar',
        marker = list(
          color = c(custom_colors$success, custom_colors$warning, "#ff9800", custom_colors$danger)
        ),
        hovertemplate = "Category: %{x}<br>Number of Species: %{y}<extra></extra>") %>%
  layout(
    title = "Number of Species by Survival Rate Category",
    xaxis = list(title = "Survival Rate Category"),
    yaxis = list(title = "Number of Species"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE
  )

p5
```
:::