---
title: "Data Explorer"
format: 
  html:
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(plotly)
library(reactable)
library(DT)
library(htmlwidgets)
library(crosstalk)
library(htmltools)

# Custom color palette
custom_colors <- list(
  primary = "#2E8B57",
  secondary = "#20B2AA", 
  success = "#28a745",
  warning = "#ffc107",
  danger = "#dc3545",
  info = "#17a2b8",
  light = "#f8f9fa",
  dark = "#343a40",
  gradient = c("#2E8B57", "#20B2AA", "#66CDAA", "#98FB98", "#90EE90")
)

# Helper functions
format_percent <- function(x, digits = 1) {
  paste0(round(x * 100, digits), "%")
}

format_number <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", scientific = FALSE)
}

# Load and process data
df_raw <- readRDS("myapp/df.rds")

df_processed <- df_raw %>%
  mutate(across(everything(), ~ na_if(., "---"))) %>%
  mutate(across(where(is.list), ~ sapply(., function(x) if(length(x) > 0) x[1] else NA))) %>%
  mutate(
    completed_time = ymd_hms(completed_time),
    started_time = ymd_hms(started_time)
  ) %>%
  mutate(
    date = as.Date(completed_time),
    week = floor_date(date, "week"),
    month = floor_date(date, "month"),
    hour_started = hour(started_time),
    day_of_week = wday(date, label = TRUE, week_start = 1),
    is_weekend = day_of_week %in% c("Sat", "Sun"),
    duration_minutes = as.numeric(difftime(completed_time, started_time, units = "mins")),
    is_night_survey = hour_started >= 19 | hour_started < 6,
    is_short_survey = duration_minutes <= 5,
    is_long_survey = duration_minutes >= 60
  ) %>%
  mutate(across(c(starts_with("ps_num_planted_"), starts_with("num_surv_")), as.numeric))

df_completed <- df_processed %>% filter(consent == 1)

# Prepare data for exploration
explorer_data <- df_completed %>%
  select(
    Date = date,
    Enumerator = username,
    Site = site,
    Woreda = woreda,
    `Duration (min)` = duration_minutes,
    `HH Size` = hh_size,
    Education = education_level,
    Age = age,
    Sex = sex,
    `Day of Week` = day_of_week,
    `Hour Started` = hour_started,
    `Is Weekend` = is_weekend,
    `Is Night Survey` = is_night_survey,
    `Is Short Survey` = is_short_survey
  ) %>%
  arrange(desc(Date))

# Create shared data for crosstalk
shared_explorer_data <- SharedData$new(explorer_data)
```

::: {.filter-section}
## ðŸ” Advanced Data Filters

::: {.interactive-controls}
```{r explorer_filters}
#| echo: false

# Date filter
filter_slider("Date", "Date Range", shared_explorer_data, ~Date, step = 1)

# Site filter
filter_select("Site", "Site", shared_explorer_data, ~Site, multiple = TRUE)

# Woreda filter
filter_select("Woreda", "Woreda", shared_explorer_data, ~Woreda, multiple = TRUE)

# Enumerator filter
filter_select("Enumerator", "Enumerator", shared_explorer_data, ~Enumerator, multiple = TRUE)

# Duration filter
filter_slider("Duration (min)", "Duration (minutes)", shared_explorer_data, ~`Duration (min)`)

# Education filter
filter_select("Education", "Education Level", shared_explorer_data, ~Education, multiple = TRUE)

# Sex filter
filter_select("Sex", "Sex", shared_explorer_data, ~Sex, multiple = TRUE)

# Day of week filter
filter_select("Day of Week", "Day of Week", shared_explorer_data, ~`Day of Week`, multiple = TRUE)

# Boolean filters
filter_checkbox("Is Weekend", "Weekend Surveys", shared_explorer_data, ~`Is Weekend`)
filter_checkbox("Is Night Survey", "Night Surveys", shared_explorer_data, ~`Is Night Survey`)
filter_checkbox("Is Short Survey", "Short Surveys (<5min)", shared_explorer_data, ~`Is Short Survey`)
```
:::
:::

## ðŸ“Š Filtered Data Summary

::: {.dashboard-grid}
::: {.metric-card}
::: {.metric-label}
Filtered Records
:::
::: {.metric-value id="filtered-count"}
Loading...
:::
:::

::: {.metric-card}
::: {.metric-label}
Unique Sites
:::
::: {.metric-value id="unique-sites"}
Loading...
:::
:::

::: {.metric-card}
::: {.metric-label}
Unique Enumerators
:::
::: {.metric-value id="unique-enumerators"}
Loading...
:::
:::

::: {.metric-card}
::: {.metric-label}
Average Duration
:::
::: {.metric-value id="avg-duration"}
Loading...
:::
:::

::: {.metric-card}
::: {.metric-label}
Date Range
:::
::: {.metric-value id="date-range"}
Loading...
:::
:::
:::

## ðŸ” Interactive Data Table

::: {.table-container}
```{r explorer_table}
#| echo: false

# Create enhanced interactive table
DT::datatable(
  shared_explorer_data,
  options = list(
    pageLength = 15,
    scrollX = TRUE,
    scrollY = "500px",
    dom = 'Bfrtip',
    buttons = list(
      'copy',
      list(extend = 'csv', filename = 'survival_survey_data'),
      list(extend = 'excel', filename = 'survival_survey_data'),
      list(extend = 'pdf', filename = 'survival_survey_data')
    ),
    columnDefs = list(
      list(className = 'dt-center', targets = c(4, 5, 6, 7, 10, 11, 12, 13, 14, 15))
    )
  ),
  extensions = c('Buttons', 'Scroller'),
  filter = 'top',
  rownames = FALSE,
  class = 'cell-border stripe hover'
) %>%
  DT::formatRound(columns = c("Duration (min)"), digits = 1) %>%
  DT::formatDate(columns = "Date", method = "toLocaleDateString") %>%
  DT::formatStyle(
    "Duration (min)",
    backgroundColor = DT::styleInterval(c(5, 30, 60), c("#ffebee", "#fff8e1", "#e8f5e8", "#ffebee"))
  ) %>%
  DT::formatStyle(
    "Is Short Survey",
    backgroundColor = DT::styleEqual(TRUE, "#ffebee")
  ) %>%
  DT::formatStyle(
    "Is Night Survey", 
    backgroundColor = DT::styleEqual(TRUE, "#fff8e1")
  ) %>%
  DT::formatStyle(
    "Is Weekend",
    backgroundColor = DT::styleEqual(TRUE, "#e3f2fd")
  )
```
:::

## ðŸ“ˆ Dynamic Visualizations

### Survey Distribution by Selected Filters

::: {.plot-container}
```{r dynamic_plots}
#| echo: false

# Time series plot
p1 <- plot_ly(shared_explorer_data, 
        x = ~Date, 
        type = 'histogram',
        marker = list(color = custom_colors$primary),
        hovertemplate = "Date: %{x}<br>Count: %{y}<extra></extra>") %>%
  layout(
    title = "Survey Distribution Over Time",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE
  )

p1
```
:::

### Duration Analysis

::: {.plot-container}
```{r duration_analysis}
#| echo: false

# Duration histogram
p2 <- plot_ly(shared_explorer_data, 
        x = ~`Duration (min)`, 
        type = 'histogram',
        nbinsx = 30,
        marker = list(color = custom_colors$secondary),
        hovertemplate = "Duration: %{x} min<br>Count: %{y}<extra></extra>") %>%
  layout(
    title = "Survey Duration Distribution",
    xaxis = list(title = "Duration (minutes)"),
    yaxis = list(title = "Frequency"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE
  )

p2
```
:::

### Site and Enumerator Distribution

::: {.plot-container}
```{r site_enum_distribution}
#| echo: false

# Site distribution
site_counts <- df_completed %>%
  count(site, sort = TRUE) %>%
  head(15)

p3 <- plot_ly(site_counts, 
        x = ~reorder(site, -n), 
        y = ~n, 
        type = 'bar',
        marker = list(color = custom_colors$gradient[3]),
        hovertemplate = "Site: %{x}<br>Surveys: %{y}<extra></extra>") %>%
  layout(
    title = "Top 15 Sites by Survey Count",
    xaxis = list(title = "Site", tickangle = -45),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p3
```
:::

### Demographic Analysis

::: {.plot-container}
```{r demographic_analysis}
#| echo: false

# Education level distribution
education_counts <- df_completed %>%
  filter(!is.na(education_level)) %>%
  count(education_level, sort = TRUE)

p4 <- plot_ly(education_counts, 
        labels = ~education_level, 
        values = ~n, 
        type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        hovertemplate = "Education: %{label}<br>Count: %{value}<br>Percentage: %{percent}<extra></extra>",
        marker = list(
          colors = custom_colors$gradient,
          line = list(color = '#FFFFFF', width = 2)
        )) %>%
  layout(
    title = "Education Level Distribution",
    paper_bgcolor = "transparent",
    showlegend = FALSE
  )

p4
```
:::

### Survey Quality Metrics

::: {.plot-container}
```{r quality_metrics}
#| echo: false

# Quality metrics by day of week
quality_by_day <- df_completed %>%
  group_by(day_of_week) %>%
  summarise(
    total_surveys = n(),
    short_surveys = sum(is_short_survey, na.rm = TRUE),
    night_surveys = sum(is_night_survey, na.rm = TRUE),
    avg_duration = mean(duration_minutes, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    short_rate = (short_surveys / total_surveys) * 100,
    night_rate = (night_surveys / total_surveys) * 100
  )

p5 <- plot_ly(quality_by_day, x = ~day_of_week, y = ~short_rate, 
        type = 'bar', name = 'Short Survey Rate (%)',
        marker = list(color = custom_colors$danger),
        hovertemplate = "Day: %{x}<br>Short Survey Rate: %{y:.1f}%<extra></extra>") %>%
  add_trace(y = ~night_rate, name = 'Night Survey Rate (%)',
            marker = list(color = custom_colors$warning),
            hovertemplate = "Day: %{x}<br>Night Survey Rate: %{y:.1f}%<extra></extra>") %>%
  layout(
    title = "Survey Quality Metrics by Day of Week",
    xaxis = list(title = "Day of Week"),
    yaxis = list(title = "Rate (%)"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    barmode = 'group',
    legend = list(orientation = 'h', y = 1.1, x = 0.5, xanchor = 'center')
  )

p5
```
:::

## ðŸ“Š Cross-tabulation Analysis

::: {.table-container}
```{r crosstab_analysis}
#| echo: false

# Create cross-tabulation of Site vs Education Level
crosstab_data <- df_completed %>%
  filter(!is.na(education_level) & !is.na(site)) %>%
  count(site, education_level) %>%
  pivot_wider(names_from = education_level, values_from = n, values_fill = 0) %>%
  arrange(desc(rowSums(select(., -site))))

reactable(
  crosstab_data,
  columns = list(
    site = colDef(
      name = "Site", 
      minWidth = 150,
      cell = function(value) {
        div(style = "font-weight: bold; color: #2E8B57;", value)
      }
    )
  ),
  defaultPageSize = 10,
  searchable = TRUE,
  highlight = TRUE,
  bordered = TRUE,
  striped = TRUE,
  theme = reactableTheme(
    headerStyle = list(backgroundColor = "#f8f9fa", fontWeight = "bold"),
    cellStyle = list(fontSize = "0.9em")
  )
)
```
:::

<script>
// Update summary statistics based on filtered data
function updateSummaryStats() {
  // This would be updated by crosstalk when filters change
  // For now, showing static values
  document.getElementById('filtered-count').textContent = '${nrow(explorer_data)}';
  document.getElementById('unique-sites').textContent = '${n_distinct(explorer_data$Site)}';
  document.getElementById('unique-enumerators').textContent = '${n_distinct(explorer_data$Enumerator)}';
  document.getElementById('avg-duration').textContent = '${round(mean(explorer_data$`Duration (min)`, na.rm = TRUE), 1)} min';
  document.getElementById('date-range').textContent = '${min(explorer_data$Date)} to ${max(explorer_data$Date)}';
}

// Initialize on page load
document.addEventListener('DOMContentLoaded', updateSummaryStats);
</script>