---
title: "Executive Summary"
format: 
  html:
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(httr)
library(jsonlite)
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(plotly)
library(reactable)
library(DT)
library(htmlwidgets)
library(htmltools)

# Custom color palette
custom_colors <- list(
  primary = "#2E8B57",
  secondary = "#20B2AA", 
  success = "#28a745",
  warning = "#ffc107",
  danger = "#dc3545",
  info = "#17a2b8",
  light = "#f8f9fa",
  dark = "#343a40",
  gradient = c("#2E8B57", "#20B2AA", "#66CDAA", "#98FB98", "#90EE90")
)

# Helper functions
format_percent <- function(x, digits = 1) {
  paste0(round(x * 100, digits), "%")
}

format_number <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", scientific = FALSE)
}

# Try to load data from CommCare API, fallback to local file if needed
tryCatch({
  # Configuration
  USERNAME <- "mohammed.seidhussen@oneacrefund.org"
  API_KEY <- "a749d18804539c5a2210817cda29630391a088bd"
  PROJECT_SPACE <- "oaf-ethiopia"
  FORM_ID <- "e24ab639e5b7d1b609cf2894f7057b75"
  
  # API Endpoint
  url <- paste0("https://www.commcarehq.org/a/", PROJECT_SPACE, "/api/v0.5/odata/forms/", FORM_ID, "/feed")
  
  # Enhanced data fetching with progress tracking
  limit <- 2000
  offset <- 0
  all_records <- list()
  
  cat("Fetching data from CommCare...\n")
  
  while (TRUE) {
    # Set query parameters
    query <- list(
      limit = limit,
      offset = offset
    )
    
    # Make API request with timeout
    response <- GET(
      url,
      query = query,
      authenticate(USERNAME, API_KEY, type = "basic"),
      timeout(10)
    )
    
    # Check response
    if (status_code(response) != 200) {
      cat(paste0("Error: ", status_code(response), "\n"))
      cat(content(response, "text"), "\n")
      break
    }
    
    # Parse response
    data <- fromJSON(content(response, "text"))
    records <- data$value
    
    if (length(records) == 0) {
      break
    }
    
    # Add records to collection
    all_records <- c(all_records, records)
    
    # Check if we have all records
    if (length(records) < limit) {
      break
    }
    
    # Update offset for next page
    offset <- offset + limit
    cat(paste0("Fetched ", length(all_records), " records so far...\n"))
  }
  
  # Convert to data frame
  df_raw <- bind_rows(all_records)
  
  # Drop farmer name and phone number 
  df_raw <- df_raw %>% 
    select(-any_of(c("farmer_name", "name", "phone_no", "tno")))
  
  cat(paste0("Successfully fetched ", nrow(df_raw), " records from CommCare\n"))
  
}, error = function(e) {
  cat("Failed to fetch from CommCare API, using local data file...\n")
  cat("Error:", e$message, "\n")
  
  # Fallback to local file
  if (file.exists("myapp/df.rds")) {
    df_raw <<- readRDS("myapp/df.rds")
    cat(paste0("Loaded ", nrow(df_raw), " records from local file\n"))
  } else {
    stop("No data source available. Please ensure either CommCare API is accessible or df.rds file exists.")
  }
})

# Process the data
df_processed <- df_raw %>%
  mutate(across(everything(), ~ na_if(., "---"))) %>%
  mutate(across(where(is.list), ~ sapply(., function(x) if(length(x) > 0) x[1] else NA))) %>%
  mutate(
    completed_time = ymd_hms(completed_time),
    started_time = ymd_hms(started_time)
  ) %>%
  mutate(
    date = as.Date(completed_time),
    week = floor_date(date, "week"),
    month = floor_date(date, "month"),
    hour_started = hour(started_time),
    day_of_week = wday(date, label = TRUE, week_start = 1),
    is_weekend = day_of_week %in% c("Sat", "Sun"),
    duration_minutes = as.numeric(difftime(completed_time, started_time, units = "mins")),
    is_night_survey = hour_started >= 19 | hour_started < 6,
    is_short_survey = duration_minutes <= 5,
    is_long_survey = duration_minutes >= 60
  ) %>%
  mutate(across(c(starts_with("ps_num_planted_"), starts_with("num_surv_")), as.numeric))

df_completed <- df_processed %>% filter(consent == 1)
```

## 📊 Key Performance Indicators

::: {.dashboard-grid}
::: {.metric-card}
::: {.metric-label}
Total Surveys
:::
::: {.metric-value}
`r format_number(nrow(df_processed))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Completed Surveys
:::
::: {.metric-value}
`r format_number(nrow(df_completed))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Refusal Rate
:::
::: {.metric-value}
`r format_percent(sum(df_processed$consent == 0, na.rm = TRUE) / nrow(df_processed))`
:::
:::

::: {.metric-card}
::: {.metric-label}
Avg. Duration
:::
::: {.metric-value}
`r round(mean(df_completed$duration_minutes, na.rm = TRUE), 1)` min
:::
:::

::: {.metric-card}
::: {.metric-label}
Unique Enumerators
:::
::: {.metric-value}
`r n_distinct(df_completed$username)`
:::
:::
:::

## 📈 Daily Survey Progress

::: {.plot-container}
```{r daily_progress}
#| echo: false

daily_data <- df_completed %>%
  group_by(date) %>%
  summarise(surveys = n(), .groups = "drop")

p1 <- plot_ly(daily_data, x = ~date, y = ~surveys, type = 'bar',
        marker = list(color = custom_colors$primary),
        hovertemplate = "Date: %{x}<br>Surveys: %{y}<extra></extra>") %>%
  layout(
    title = "Daily Survey Completion",
    xaxis = list(title = "Date"),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE
  )

p1
```
:::

## 🎯 Target Achievement

::: {.plot-container}
```{r gauge_plot}
#| echo: false

target <- 3600
current <- nrow(df_completed)
percentage <- min(current / target, 1) * 100

gauge_plot <- plot_ly(
  type = "indicator",
  mode = "gauge+number+delta",
  value = current,
  delta = list(reference = target),
  gauge = list(
    axis = list(range = list(NULL, target)),
    bar = list(color = custom_colors$primary),
    steps = list(
      list(range = c(0, target*0.5), color = "lightgray"),
      list(range = c(target*0.5, target*0.8), color = "gray")
    ),
    threshold = list(
      line = list(color = "red", width = 4),
      thickness = 0.75,
      value = target
    )
  )
) %>% 
layout(
  title = "Survey Target Progress", 
  paper_bgcolor = "transparent",
  font = list(color = custom_colors$dark)
)

gauge_plot
```
:::

## 📈 Weekly Trends

::: {.plot-container}
```{r weekly_trends}
#| echo: false

weekly_data <- df_completed %>%
  group_by(week) %>%
  summarise(surveys = n(), .groups = "drop")

p3 <- plot_ly(weekly_data, x = ~week, y = ~surveys, type = 'scatter', mode = 'lines+markers',
        line = list(color = custom_colors$primary, width = 3), 
        marker = list(color = custom_colors$primary, size = 8),
        hovertemplate = "Week: %{x}<br>Surveys: %{y}<extra></extra>") %>%
  layout(
    title = "Weekly Survey Trends",
    xaxis = list(title = "Week"),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE
  )

p3
```
:::

## 🗺️ Geographic Distribution

::: {.plot-container}
```{r geographic_plot}
#| echo: false

geo_data <- df_completed %>%
  group_by(woreda) %>%
  summarise(count = n(), .groups = "drop")

p4 <- plot_ly(geo_data, x = ~reorder(woreda, -count), y = ~count, 
        type = 'bar', marker = list(color = custom_colors$secondary),
        hovertemplate = "Woreda: %{x}<br>Surveys: %{y}<extra></extra>") %>%
  layout(
    title = "Surveys by Woreda",
    xaxis = list(title = "Woreda", tickangle = -45),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 100)
  )

p4
```
:::