---
title: "Site Analysis"
format: 
  html:
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| include: false
library(dplyr)
library(tidyr)
library(lubridate)
library(stringr)
library(plotly)
library(reactable)
library(DT)
library(htmlwidgets)
library(crosstalk)
library(htmltools)

# Custom color palette
custom_colors <- list(
  primary = "#2E8B57",
  secondary = "#20B2AA", 
  success = "#28a745",
  warning = "#ffc107",
  danger = "#dc3545",
  info = "#17a2b8",
  light = "#f8f9fa",
  dark = "#343a40",
  gradient = c("#2E8B57", "#20B2AA", "#66CDAA", "#98FB98", "#90EE90")
)

# Helper functions
format_percent <- function(x, digits = 1) {
  paste0(round(x * 100, digits), "%")
}

format_number <- function(x, digits = 0) {
  format(round(x, digits), big.mark = ",", scientific = FALSE)
}

# Load and process data
df_raw <- readRDS("myapp/df.rds")

df_processed <- df_raw %>%
  mutate(across(everything(), ~ na_if(., "---"))) %>%
  mutate(across(where(is.list), ~ sapply(., function(x) if(length(x) > 0) x[1] else NA))) %>%
  mutate(
    completed_time = ymd_hms(completed_time),
    started_time = ymd_hms(started_time)
  ) %>%
  mutate(
    date = as.Date(completed_time),
    week = floor_date(date, "week"),
    month = floor_date(date, "month"),
    hour_started = hour(started_time),
    day_of_week = wday(date, label = TRUE, week_start = 1),
    is_weekend = day_of_week %in% c("Sat", "Sun"),
    duration_minutes = as.numeric(difftime(completed_time, started_time, units = "mins")),
    is_night_survey = hour_started >= 19 | hour_started < 6,
    is_short_survey = duration_minutes <= 5,
    is_long_survey = duration_minutes >= 60
  ) %>%
  mutate(across(c(starts_with("ps_num_planted_"), starts_with("num_surv_")), as.numeric))

df_completed <- df_processed %>% filter(consent == 1)

# Site statistics
site_stats <- df_completed %>%
  group_by(site) %>%
  summarise(
    total_surveys = n(),
    avg_duration = round(mean(duration_minutes, na.rm = TRUE), 1),
    completion_rate = round((n() / nrow(df_completed)) * 100, 1),
    unique_enumerators = n_distinct(username),
    avg_hh_size = round(mean(hh_size, na.rm = TRUE), 1),
    total_planted = sum(across(starts_with("ps_num_planted_")), na.rm = TRUE),
    total_survived = sum(across(starts_with("num_surv_")), na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    survival_rate = ifelse(total_planted > 0, (total_survived / total_planted) * 100, 0)
  ) %>%
  arrange(desc(total_surveys))

# Create shared data for crosstalk
shared_site_data <- SharedData$new(site_stats)
```

::: {.filter-section}
## 🔍 Site Analysis Filters

::: {.interactive-controls}
```{r site_filters}
#| echo: false

# Site filter
filter_select("site", "Site", shared_site_data, ~site, multiple = TRUE)

# Survey count filter
filter_slider("total_surveys", "Total Surveys", shared_site_data, ~total_surveys)

# Survival rate filter
filter_slider("survival_rate", "Survival Rate (%)", shared_site_data, ~survival_rate)

# Duration filter
filter_slider("avg_duration", "Average Duration (min)", shared_site_data, ~avg_duration)
```
:::
:::

## 🏢 Site Performance Overview

::: {.dashboard-grid}
::: {.metric-card}
::: {.metric-label}
Total Sites
:::
::: {.metric-value}
`r nrow(site_stats)`
:::
:::

::: {.metric-card}
::: {.metric-label}
Most Active Site
:::
::: {.metric-value}
`r site_stats$site[1]`
:::
:::

::: {.metric-card}
::: {.metric-label}
Best Survival Rate
:::
::: {.metric-value}
`r format_percent(max(site_stats$survival_rate, na.rm = TRUE) / 100)`
:::
:::

::: {.metric-card}
::: {.metric-label}
Average Duration
:::
::: {.metric-value}
`r round(mean(site_stats$avg_duration, na.rm = TRUE), 1)` min
:::
:::

::: {.metric-card}
::: {.metric-label}
Total Enumerators
:::
::: {.metric-value}
`r sum(site_stats$unique_enumerators, na.rm = TRUE)`
:::
:::
:::

## 🏢 Site Performance Comparison

::: {.plot-container}
```{r site_performance_plot}
#| echo: false

p1 <- plot_ly(site_stats, 
        x = ~reorder(site, -total_surveys), 
        y = ~total_surveys, 
        type = 'bar', 
        marker = list(color = custom_colors$gradient[1]),
        hovertemplate = "Site: %{x}<br>Surveys: %{y}<br>Avg Duration: %{customdata[0]} min<br>Enumerators: %{customdata[1]}<extra></extra>",
        customdata = ~cbind(avg_duration, unique_enumerators)) %>%
  layout(
    title = "Survey Count by Site",
    xaxis = list(title = "Site", tickangle = -45),
    yaxis = list(title = "Number of Surveys"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p1
```
:::

## 📊 Comprehensive Site Comparison Table

::: {.table-container}
```{r site_comparison_table}
#| echo: false

reactable(
  shared_site_data,
  columns = list(
    site = colDef(
      name = "Site", 
      minWidth = 150,
      cell = function(value) {
        div(style = "font-weight: bold; color: #2E8B57;", value)
      }
    ),
    total_surveys = colDef(
      name = "Total Surveys", 
      align = "center", 
      minWidth = 100,
      cell = function(value) {
        color <- if (value >= 100) "#28a745" else if (value >= 50) "#ffc107" else "#dc3545"
        div(style = paste("color:", color, "; font-weight: bold;"), format(value, big.mark = ","))
      }
    ),
    avg_duration = colDef(
      name = "Avg Duration (min)", 
      align = "center",
      format = colFormat(digits = 1),
      cell = function(value) {
        color <- if (value <= 20) "#28a745" else if (value <= 40) "#ffc107" else "#dc3545"
        div(style = paste("color:", color, "; font-weight: bold;"), paste(value, "min"))
      }
    ),
    completion_rate = colDef(
      name = "Share of Total (%)", 
      align = "center",
      format = colFormat(digits = 1),
      cell = function(value) {
        div(style = "font-weight: bold;", paste0(value, "%"))
      }
    ),
    unique_enumerators = colDef(
      name = "Enumerators", 
      align = "center"
    ),
    avg_hh_size = colDef(
      name = "Avg HH Size", 
      align = "center",
      format = colFormat(digits = 1)
    ),
    total_planted = colDef(
      name = "Trees Planted", 
      align = "center",
      format = colFormat(separators = TRUE)
    ),
    total_survived = colDef(
      name = "Trees Survived", 
      align = "center",
      format = colFormat(separators = TRUE)
    ),
    survival_rate = colDef(
      name = "Survival Rate (%)",
      align = "center",
      format = colFormat(digits = 1),
      cell = function(value) {
        color <- if (value >= 80) {
          "#A5D6A7"
        } else if (value >= 60) {
          "#FFF9C4"
        } else {
          "#FFCDD2"
        }
        div(
          style = paste("background:", color, "; padding: 4px; border-radius: 4px; font-weight: bold;"),
          paste0(round(value, 1), "%")
        )
      }
    )
  ),
  defaultPageSize = 10, 
  searchable = TRUE, 
  highlight = TRUE, 
  bordered = TRUE, 
  striped = TRUE,
  theme = reactableTheme(
    headerStyle = list(backgroundColor = "#f8f9fa", fontWeight = "bold"),
    cellStyle = list(fontSize = "0.9em")
  )
)
```
:::

## 🌱 Tree Survival Rates by Site

::: {.plot-container}
```{r site_survival_plot}
#| echo: false

p2 <- plot_ly(site_stats, 
        x = ~reorder(site, survival_rate), 
        y = ~survival_rate, 
        type = 'bar', 
        marker = list(
          color = ~survival_rate, 
          colorscale = list(c(0, "#ffcdd2"), c(1, custom_colors$success)),
          showscale = TRUE,
          colorbar = list(title = "Survival Rate (%)")
        ),
        hovertemplate = "Site: %{x}<br>Planted: %{customdata[0]}<br>Survived: %{customdata[1]}<br>Rate: %{y:.1f}%<extra></extra>",
        customdata = ~cbind(total_planted, total_survived)) %>%
  layout(
    title = "Tree Survival Rates by Site",
    xaxis = list(title = "Site", tickangle = -45),
    yaxis = list(title = "Survival Rate (%)", range = c(0, 100)),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p2
```
:::

## ⏱️ Average Survey Duration by Site

::: {.plot-container}
```{r site_duration_plot}
#| echo: false

duration_data <- site_stats %>%
  arrange(desc(avg_duration))

p3 <- plot_ly(duration_data, 
        x = ~reorder(site, avg_duration), 
        y = ~avg_duration, 
        type = 'bar', 
        marker = list(color = custom_colors$warning),
        hovertemplate = "Site: %{x}<br>Avg Duration: %{y} min<br>Total Surveys: %{customdata}<extra></extra>",
        customdata = ~total_surveys) %>%
  layout(
    title = "Average Survey Duration by Site",
    xaxis = list(title = "Site", tickangle = -45),
    yaxis = list(title = "Average Duration (minutes)"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p3
```
:::

## 📈 Site Performance Matrix

::: {.plot-container}
```{r site_performance_matrix}
#| echo: false

p4 <- plot_ly(site_stats, 
        x = ~total_surveys, 
        y = ~survival_rate, 
        type = 'scatter',
        mode = 'markers',
        marker = list(
          size = ~avg_duration,
          sizemode = 'diameter',
          sizeref = 2 * max(site_stats$avg_duration) / (40^2),
          color = ~unique_enumerators,
          colorscale = 'Viridis',
          showscale = TRUE,
          colorbar = list(title = "Enumerators")
        ),
        text = ~site,
        hovertemplate = "Site: %{text}<br>Surveys: %{x}<br>Survival Rate: %{y:.1f}%<br>Avg Duration: %{marker.size} min<br>Enumerators: %{marker.color}<extra></extra>") %>%
  layout(
    title = "Site Performance Matrix (Size = Duration, Color = Enumerators)",
    xaxis = list(title = "Total Surveys"),
    yaxis = list(title = "Survival Rate (%)"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent"
  )

p4
```
:::

## 🏆 Site Rankings

::: {.plot-container}
```{r site_rankings}
#| echo: false

# Create rankings for different metrics
site_rankings <- site_stats %>%
  mutate(
    survey_rank = rank(-total_surveys),
    survival_rank = rank(-survival_rate),
    duration_rank = rank(avg_duration),  # Lower duration is better
    overall_score = (survey_rank + survival_rank + duration_rank) / 3
  ) %>%
  arrange(overall_score) %>%
  select(site, survey_rank, survival_rank, duration_rank, overall_score) %>%
  head(10)

# Reshape for plotting
rankings_long <- site_rankings %>%
  pivot_longer(cols = c(survey_rank, survival_rank, duration_rank), 
               names_to = "metric", values_to = "rank") %>%
  mutate(
    metric = case_when(
      metric == "survey_rank" ~ "Survey Count",
      metric == "survival_rank" ~ "Survival Rate", 
      metric == "duration_rank" ~ "Duration (Lower=Better)"
    )
  )

p5 <- plot_ly(rankings_long, 
        x = ~metric, 
        y = ~rank, 
        color = ~site,
        type = 'scatter',
        mode = 'lines+markers',
        line = list(width = 3),
        marker = list(size = 8),
        hovertemplate = "Site: %{fullData.name}<br>Metric: %{x}<br>Rank: %{y}<extra></extra>") %>%
  layout(
    title = "Site Rankings Across Key Metrics (Lower Rank = Better Performance)",
    xaxis = list(title = "Performance Metric"),
    yaxis = list(title = "Rank", autorange = "reversed"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    legend = list(orientation = 'h', y = 1.15, x = 0.5, xanchor = 'center')
  )

p5
```
:::

## 📊 Site Efficiency Analysis

::: {.plot-container}
```{r site_efficiency}
#| echo: false

# Calculate efficiency metrics
efficiency_data <- site_stats %>%
  mutate(
    surveys_per_enumerator = total_surveys / unique_enumerators,
    trees_per_survey = total_planted / total_surveys,
    efficiency_score = (surveys_per_enumerator * survival_rate) / avg_duration
  ) %>%
  arrange(desc(efficiency_score)) %>%
  head(10)

p6 <- plot_ly(efficiency_data, 
        x = ~reorder(site, efficiency_score), 
        y = ~efficiency_score, 
        type = 'bar',
        marker = list(color = custom_colors$info),
        hovertemplate = "Site: %{x}<br>Efficiency Score: %{y:.2f}<br>Surveys/Enumerator: %{customdata[0]:.1f}<br>Trees/Survey: %{customdata[1]:.1f}<extra></extra>",
        customdata = ~cbind(surveys_per_enumerator, trees_per_survey)) %>%
  layout(
    title = "Site Efficiency Score (Top 10)",
    xaxis = list(title = "Site", tickangle = -45),
    yaxis = list(title = "Efficiency Score"),
    paper_bgcolor = "transparent", 
    plot_bgcolor = "transparent",
    showlegend = FALSE,
    margin = list(b = 120)
  )

p6
```
:::